To log or not to log

### 2022-08-21 main.tex
manuscript update

As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}. Every week the European Forecast Hub collates and aggregates forecasts for different COVID-19 related targets from teams around the world. Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ... 0.95, 0.975, 0.99$). The forecasts for the purpose of this illustrations are two-week-ahead forecasts made between XXX and XXX for reported cases and deaths from COVID-19 in XXX countries. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates. Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

    \begin{tabular}{lllcc}
    \toprule
    target\_type & quantity & measure & natural & log\\
    \midrule
    Cases & Observations & mean & 20308 & 8.39\\
    Cases & Observations & sd & 43405 & 2.11\\
    Cases & Observations & var & 1884002308 & 4.46\\
    \addlinespace
    Deaths & Observations & mean & 259 & 3.79\\
    Deaths & Observations & sd & 532 & 2.11\\
    Deaths & Observations & var & 283492 & 4.46\\
    \addlinespace
    \hline
    \addlinespace
    Cases & WIS & mean & 4544 & 0.28\\
    Cases & WIS & sd & 17585 & 0.49\\
    \addlinespace
    Deaths & WIS & mean & 28 & 0.22\\
    Deaths & WIS & sd & 67 & 0.26\\
    \bottomrule
    \end{tabular}

    \caption{Table with summary statistics for observations and scores for forecasts from the ECDC data set.}
    \label{tab:HUB-summary}
\end{table}

Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}). The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B). Observations are overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}). 
%maybe make a line of best fit of the form variance = alpha * mean^2 + beta * mean

@@

Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}). On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).

Variation of scores across targets (see Figure \ref{fig:HUBXXX} and locations (see Figure \ref{fig:HUB-mean-locations}A, B) is much larger on the natural than on the log scale, where scores are more evenly distributed. On the natural scale, average scores per location correlate strongly (XX) with the average number of observed cases or deaths (See Figure XX). On the log scale, this correlation is much weaker and slightly negative. Forecasts across different locations receive more similar scores, with a tendency for scores to be slightly higher in locations with few observed cases and deaths (see Figures XX and \ref{fig:HUB-mean-locations}. 
%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this. 
For individual forecasts, rankings between models are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}).

@@

@@

When evaluated on the natural scale, there is  strong relationship between the mean of the weighted interval score and the total number of observed cases (correlation: 0.913) or deaths (cor: 0.849) in a location (see Figure \ref{fig:HUB-mean-scores-total-loglog} and Figure \ref{fig:HUB-mean-scores-total} in the SI). The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
    \caption{Plot with Weighted interval scores against the mean number of observed cases or deaths.}
    \label{fig:HUB-mean-scores-total-loglog}
\end{figure}

@@

\paragraph{Rankings}
So one effect is that for a given dispersion on the natural scale, it's better if that interval is skewed, as that implies that at least you only off a lot in one direction in negative terms, and not in the other direction. Maybe looking at over-and underprediction is the key here. 
%ADD SAME PLOT TO SI FOR HORIZON 4.

\paragraph{pairwise comparisons}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
    \caption{Changes in relative skill and the distribution of scores}
    \label{fig:HUB-pairwise}
\end{figure}

Correlation between pairwise comparisons is 0.4414717.

@@

>>>>

\paragraph{Intro Hub}
As an empirical example for evaluating forecasts on the natural and on the log scale we use forecasts from the European Forecast Hub \citep{europeancovid-19forecasthubEuropeanCovid19Forecast2021, sherrattPredictivePerformanceMultimodel2022}. Every week the European Forecast Hub collates and aggregates forecasts for different COVID-19 related targets from teams around the world. Forecasts are made one to four weeks ahead into the future and follow a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ... 0.95, 0.975, 0.99$). The forecasts for the purpose of this illustrations are two-week-ahead forecasts made between XXX and XXX for reported cases and deaths from COVID-19 in XXX countries. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 16 forecast dates. Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

In addition to the WIS we use pairwise comparisons (CITATION Kramer) to evaluate models against each other. In a first step, score ratios are computed for all pairs of models by taking the set of overlapping forecasts between the two models and dividing the score of one model over the score achieved by the other model. The relative skill for a given model is then obtained by taking the geometric mean of all score ratios which involve that model. Low values are better, and the average model receives a relative skill score of 1.

@@

SECTION STRUCUTRE / IMPORTANT THINGS
\begin{itemize}
\item X variation gets smaller / more even
\item X relationship with country size changes
\item X inndividual rankings stay similar
\item relative skill scores change a lot
\item deep dive into what happens for individual models
\item scores increase across forecast horizons
\end{itemize}

Maybe look at correlation between bias and score?
Correlation between relative WIS contributions.?

\paragraph{Mean / sd of cases and WIS values}
Pooled across all time points and locations, the mean number of cases (deaths) observed was 22811 (368) with an overall standard deviation of 47135 (831) (see Table \ref{tab:HUB-summary}). The average number of observed cases and deaths varies considerably by location (see Figure \ref{fig:HUB-mean-locations}A and B). Observations were overdispersed, meaning that for a given location, the variance in observations exceeds the mean of all observed values (see Figure \ref{fig:HUB-variance-mean}). 
%Could fit a model var = alpha * x2 + beta *x

On the natural scale, scores vary considerably across targets (see Fig \ref{}) and locations (see Fig \ref{}) and correlate strongly (XX) with the average number of observed cases / deaths in a given country (see Fig \ref{fig:HUB-mean-locations}A, B). On the log scale, scores are more evenly distributed between targets (see Fig \ref{}) and locations (see Fig \ref{fig:HUB-mean-locations}. The correlation between average WIS and the mean number of observations in a country is much weaker and tends to be negative. Locations therefore tend to have a more equal influence on avearage scores, with smaller countries potentially being slightly overrepresented. 
%This is in line with the observation that the association between variance is smaller than mu^2 + mu, meaning that we would expect this.

 % Pooled across all time points and locations, the mean WIS on the natural scale for cases (deaths) was 8855 (59.3) with an overall standard deviation of 51274 (156) (see Table \ref{tab:HUB-summary}). On the log scale, scores were much closer to each other (see Figure \ref{fig:HUB-average-scores}), with a mean WIS for cases (deaths) of 0.507 (0.362) with an overall standard deviation of 0.745 (0.407).

% When evaluated on the natural scale, there is  strong relationship between the mean of the weighted interval score and the total number of observed cases (correlation: 0.913) or deaths (cor: 0.849) in a location (see Figure \ref{fig:HUB-mean-scores-total-loglog} and Figure \ref{fig:HUB-mean-scores-total} in the SI). The relationship is less pronounced for scores on the log scale (cor: -0.019 for cases, -0.395 for deaths) and negative, meaning that locations with fewer observed cases or deaths tend to receive higher scores.

@@

\paragraph{individual rankings}
For individual forecasts, rankings between models for single forecasts are mostly preserved (correlation: X for cases, Y for deaths), with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}). When evaluating performance across different forecasts and forecast targets, relative skill scores of the models change considerably. At least for cases, the effect is stronger with increasing forecast horizon.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-mean-scores-vs-total-log-log.png}
    \caption{My current thinking is that the top row is helpful, as it explalins what the mean / var relationship is which is relevant for how much weight small countries get. I think the 2nd row is just completely useless. The 3rd and 4th row may be helpful. In some sense it is duplicated information from above, but it gives some nice intuiton for the correlations. I could alternatively also turn it into a correlation plot and add it to the other one. }
    \label{fig:HUB-mean-scores-total-loglog}
\end{figure}

@@

% \paragraph{Rankings}
% So one effect is that for a given dispersion on the natural scale, it's better if that interval is skewed, as that implies that at least you only off a lot in one direction in negative terms, and not in the other direction. Maybe looking at over-and underprediction is the key here. 
%ADD SAME PLOT TO SI FOR HORIZON 4.

\paragraph{pairwise comparisons}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{output/figures/HUB-pairwise-comparisons.png}
    \caption{Changes in relative skill and the distribution of scores. MAYBE THIS SHOULD ONLY BE CASES?}
    \label{fig:HUB-pairwise}
\end{figure}

% Correlation between pairwise comparisons is 0.4414717.

@@

    \begin{tabular}{lllcc}
    \toprule
    target\_type & quantity & measure & natural & log\\
    \midrule
    Cases & Observations & mean & 20308 & 8.39\\
    Cases & Observations & sd & 43405 & 2.11\\
    Cases & Observations & var & 1884002308 & 4.46\\
    \addlinespace
    Deaths & Observations & mean & 259 & 3.79\\
    Deaths & Observations & sd & 532 & 2.11\\
    Deaths & Observations & var & 283492 & 4.46\\
    \addlinespace
    \hline
    \addlinespace
    Cases & WIS & mean & 4544 & 0.28\\
    Cases & WIS & sd & 17585 & 0.49\\
    \addlinespace
    Deaths & WIS & mean & 28 & 0.22\\
    Deaths & WIS & sd & 67 & 0.26\\
    \bottomrule
    \end{tabular}
    \caption{Table with summary statistics for observations and scores for forecasts from the ECDC data set.}
    \label{tab:HUB-summary}
\end{table}
